11th International Satisfiability Modulo Theories Competition (SMT-COMP 2016): Rules and Procedures

### 2016-04-22 rules/rules16.tex
Unsat core track.

SMT-COMP 2016 will have two tracks: the conventional main track and an
application (i.e., incremental) track.  Within each track there are
multiple divisions, where each division uses benchmarks from a
specific SMT-LIB logic (or group of logics).  We will recognize
winners as measured by number of benchmarks solved; we will also
recognize solvers based on additional criteria.

The rest of this document, revised from the previous
version,\footnote{Earlier versions of this document include
  contributions from Clark Barrett, Roberto Bruttomesso, David Cok,
  Sylvain Conchon, David D{\'e}harbe, Morgan Deters, Alberto Griggio,
  Albert Oliveras, Aaron Stump, and Tjark Weber.} describes the rules
and competition procedures for SMT-COMP~2016.  The principal changes
from the previous competition rules are the following:
\begin{itemize}
\item Benchmarks will use (a subset of) version~2.5 of the SMT-LIB
  language.  \emph{Rationale:} This is the latest version of the
  SMT-LIB language.  It was released on 2015-06-28, and is largely
  backwards-compatible to version~2.0, which was used for SMT-COMP
  2015.
\item Solver output may affect the score even when the solver does not
  terminate within the time limit.  (In 2015, main track solver output
  was ignored if the solver subsequently timed out.)  Solvers should
  take care not to accidentally produce output that contains
  \texttt{sat} or \texttt{unsat} even when they are killed.
  \emph{Rationale:} Users are likely to trust solver responses even
  when the solver continues to run for some time.
\item Divisions for floating-point arithmetic are no longer
  experimental, and will be considered competitive if the necessary
  requirements (see Section~\ref{sec:scoring}) are met.
  \emph{Rationale:} Floating-point divisions were experimental in
  2015.  By now, their definition is sufficiently stable, and they are
  supported by several solvers.
\item Division scores will be based on a weighted sum of scores for
  benchmark families.  \emph{Rationale:} For some years now, SMT-COMPs
  have had sufficient computational resources to evaluate all solver
  entrants on all eligible benchmarks.  As a side-effect, the
  weighting of benchmark families that was achieved in early SMT-COMPs
  through a careful selection of benchmarks (based on benchmark
  difficulty and category) was lost.  Since there are vast differences
  in size between benchmark families, we believe that a weighting that
  de-emphasizes large benchmark families (see
  Section~\ref{sec:division-scoring}) will lead to more meaningful
  competition results.
\item Best industrial performance will no longer be recognized
  separately.  \emph{Rationale:} While there is agreement in the SMT
  community to emphasize problems that come from real applications,
  industrial performance largely coincided with overall performance in
  SMT-COMP 2015.  We consider the effort to determine and report it
  separately no longer justified.
\end{itemize}

@@

\header{Input and output.}
%
In the main track, a participating solver must read a single benchmark
script, whose filename is presented as the solver's first command-line
argument.  In the application track, a trace executor will send
commands from a benchmark script to the solver's standard input
channel.

The benchmark script is in the concrete syntax of the SMT-LIB format,
version~2.5, though with a restricted set of commands.  A benchmark
script is a text file containing a sequence of SMT-LIB commands that
satisfies the following requirements:
%
\begin{enumerate}
\item In the main track, there may be a (single) \akey{set-option
  :print-success ...} command.  Note that \texttt{success} outputs are
  ignored by the post-processor used by the
  competition.\footnote{SMT-LIB~2.5 requires solvers to produce a
    \texttt{success} answer after each \akey{set-logic},
    \akey{declare-sort}, \akey{declare-fun} and \akey{assert} command
    (among others), unless the option \akey{:print-success} is set to
    false.  Ignoring the \texttt{success} outputs allows for
    submitting fully SMT-LIB~2.5 compliant solvers without the need
    for a wrapper script, while still allowing entrants of previous
    competitions to run without changes.}
\item In the application track, the \akey{:print-success} option must
  not be disabled.  The trace executor will send an initial
  $\akey{set-option :print-success true}$ command to the solver.
\item The (single) $\akey{set-logic}$ command setting the benchmark's
  logic is the first command after any $\akey{set-option}$ commands.
\item The script may contain any number of $\akey{set-info}$ commands.
  These may be ignored by solvers.
\item The script may contain any number of $\akey{declare-sort}$
  commands.  All sorts declared with a $\akey{declare-sort}$ command
  must have zero arity.
\item The script may contain any number of $\akey{define-sort}$
  commands.  All sorts defined with a $\akey{define-sort}$ command
  must have zero arity.
\item The script may contain any number of $\akey{declare-fun}$
  commands.
\item The script may contain any number of $\akey{define-fun}$
  commands.
\item The script may contain any number of $\akey{assert}$ commands.
  All formulas in the script belong to the benchmark's logic, with any
  free symbols declared in the script.
\item Named formulas are not used in benchmark scripts.
\item In the main track, there is exactly one $\akey{check-sat}$
  command.
\item In the application track, there are one or more \akey{check-sat}
  commands.  There may also be zero or more \akey{push 1} commands,
  and zero or more \akey{pop 1} commands, consistent with the use of
  those commands in the SMT-LIB standard.
\item The script may optionally contain an $\akey{exit}$ command as
  its last command.  In the application track, this command must not
  be omitted.
\item No other commands besides the ones just mentioned may be used.
\end{enumerate}
%
The SMT-LIB format specification is available from the ``Standard''
section of the SMT-LIB website~\cite{SMT-LIB}.  Solvers will be given
formulas only from the divisions into which they have been entered.

@@

@@

\header{Benchmark sources.}
%
Benchmarks for each division will be drawn from the SMT-LIB benchmark
library.  The main track will use a subset of all
\emph{non-incremental} benchmarks; the application track will use a
subset of all \emph{incremental} benchmarks.

@@

\header{Main track.} More specifically, for the main track, we have
%
\begin{itemize}
\item $e=0$, $n=0$ if the solver aborts without a response, or the
  result of the \akey{check-sat} command is \texttt{unknown},
\item $e=0$, $n=1$ if the result of the \akey{check-sat} command is
  correct,
\item $e=1$, $n=0$ if the result of the \akey{check-sat} command is
  incorrect.
\end{itemize}
%
Note that a (correct or incorrect) response is taken into
consideration even when the the solver process terminates abnormally,
or does not terminate within the time limit.  Solvers should take care
not to accidentally produce output that contains \texttt{sat} or
\texttt{unsat}.

@@

@@

For a given competition benchmark~$b$, let~$F_b \geq 1$ be the total
number of benchmarks in $b$'s benchmark family that were used in the
competition.  We define the weight for benchmark~$b$ as $\alpha_b = (1
+ \log_e F_b) / F_b$.\footnote{See Section~\ref{sec:competition-wide}
  for a motivating discussion of log scaling.}  We define the
\emph{normalized weight} for benchmark~$b$ as $\alpha'_b = \alpha_b /
(\sum_{b'} \alpha_{b'})$, where the sum is over all benchmarks in the
division.

For main track divisions, we will separately compute the weighted sum
of all raw scores (Section~\ref{sec:benchmark-scoring}) $$\sum_b
\alpha'_b \cdot \langle e_b, n_b, w_b, c_b\rangle$$ where the sum is
over all benchmarks in the division to assess parallel performance,
and the weighted sum of all sequential scores
(Section~\ref{sec:sequential}) to assess sequential performance.  For
application track divisions, division scores will be based on raw
scores only.\footnote{Since application track benchmarks may be
  partially solved, defining a useful sequential score for the
  application track would require information not provided by the raw
  score, e.g., detailed timing information for each result.}

>>>>

SMT-COMP 2016 will have three tracks: the conventional main track, an
application (i.e., incremental) track, and an unsat-core track.  The
unsat-core track will be experimental in 2016.

Within each track there are multiple divisions, where each division
uses benchmarks from a specific SMT-LIB logic (or group of logics).
We will recognize winners as measured by number of benchmarks solved;
we will also recognize solvers based on additional criteria.

The rest of this document, revised from the previous
version,\footnote{Earlier versions of this document include
  contributions from Clark Barrett, Roberto Bruttomesso, David Cok,
  Sylvain Conchon, David D{\'e}harbe, Morgan Deters, Alberto Griggio,
  Albert Oliveras, Aaron Stump, and Tjark Weber.} describes the rules
and competition procedures for SMT-COMP~2016.  The principal changes
from the previous competition rules are the following:
\begin{itemize}
\item Benchmarks will use (a subset of) version~2.5 of the SMT-LIB
  language.  \emph{Rationale:} This is the latest version of the
  SMT-LIB language.  It was released on 2015-06-28, and is largely
  backwards-compatible to version~2.0, which was used for SMT-COMP
  2015.
\item Solver output may affect the score even when the solver does not
  terminate within the time limit.  (In 2015, main track solver output
  was ignored if the solver subsequently timed out.)  Solvers should
  take care not to accidentally produce output that contains
  \texttt{sat} or \texttt{unsat} even when they are killed.
  \emph{Rationale:} Users are likely to trust solver responses even
  when the solver continues to run for some time.
\item Divisions for floating-point arithmetic are no longer
  experimental, and will be considered competitive if the necessary
  requirements (see Section~\ref{sec:scoring}) are met.
  \emph{Rationale:} Floating-point divisions were experimental in
  2015.  By now, their definition is sufficiently stable, and they are
  supported by several solvers.
\item Division scores will be based on a weighted sum of scores for
  benchmark families.  \emph{Rationale:} For some years now, SMT-COMPs
  have had sufficient computational resources to evaluate all solver
  entrants on all eligible benchmarks.  As a side-effect, the
  weighting of benchmark families that was achieved in early SMT-COMPs
  through a careful selection of benchmarks (based on benchmark
  difficulty and category) was lost.  Since there are vast differences
  in size between benchmark families, we believe that a weighting that
  de-emphasizes large benchmark families (see
  Section~\ref{sec:division-scoring}) will lead to more meaningful
  competition results.
\item The unsat-core track that was introduced in SMT-COMP 2012, but
  discontinued in 2013-2015 (partly because of the competition's move
  to StarExec), is back.  It will be experimental in 2016; its results
  will be reported, but no official winners will be announced.
  \emph{Rationale:} Unsat cores are important in many applications of
  SMT solving.  The track was discontinued primarily for
  infrastructure and resource reasons.  We are pleased that we can
  finally offer it again, but we ask for your understanding in case of
  mishaps while we port the required tools.
\item Best industrial performance will no longer be recognized
  separately.  \emph{Rationale:} While there is agreement in the SMT
  community to emphasize problems that come from real applications,
  industrial performance largely coincided with overall performance in
  SMT-COMP 2015.  We consider the effort to determine and report it
  separately no longer justified.
\end{itemize}

@@

\header{Input and output.}
%
In the main and unsat-core track, a participating solver must read a
single benchmark script, whose filename is presented as the solver's
first command-line argument.  In the application track, a trace
executor will send commands from a benchmark script to the solver's
standard input channel.

The benchmark script is in the concrete syntax of the SMT-LIB format,
version~2.5, though with a restricted set of commands.  A benchmark
script is a text file containing a sequence of SMT-LIB commands that
satisfies the following requirements:
%
\begin{enumerate}
\item
  \begin{enumerate}
  \item In the main and unsat-core track, there may be a single
    \akey{set-option :print-success ...} command.  Note that
    \texttt{success} outputs are ignored by the post-processor used by
    the competition.\footnote{SMT-LIB~2.5 requires solvers to produce
      a \texttt{success} answer after each \akey{set-logic},
      \akey{declare-sort}, \akey{declare-fun} and \akey{assert}
      command (among others), unless the option \akey{:print-success}
      is set to false.  Ignoring the \texttt{success} outputs allows
      for submitting fully SMT-LIB~2.5 compliant solvers without the
      need for a wrapper script, while still allowing entrants of
      previous competitions to run without changes.}
  \item In the application track, the \akey{:print-success} option
    must not be disabled.  The trace executor will send an initial
    \akey{set-option :print-success true} command to the solver.
  \end{enumerate}
\item In the unsat-core track, there is a single \akey{set-option
  :produce-unsat-cores true} command.
\item The (single) \akey{set-logic} command setting the benchmark's
  logic is the first command after any \akey{set-option} commands.
\item The script may contain any number of \akey{set-info} commands.
\item The script may contain any number of \akey{declare-sort} and
  \akey{define-sort} commands.  All sorts declared or defined with
  these commands must have zero arity.
\item The script may contain any number of \akey{declare-fun} and
  \akey{define-fun} commands.
\item The script may contain any number of \akey{assert} commands.
  All formulas in the script belong to the benchmark's logic, with any
  free symbols declared in the script.
\item
  \begin{enumerate}
    \item In the main and application track, named formulas are not
      used.
    \item In the unsat core track, top-level assertions may be named.
  \end{enumerate}
\item
  \begin{enumerate}
  \item In the main and unsat-core track, there is exactly one
    \akey{check-sat} command.
  \item In the application track, there are one or more
    \akey{check-sat} commands.  There may also be zero or more
    \akey{push 1} commands, and zero or more \akey{pop 1} commands,
    consistent with the use of those commands in the SMT-LIB standard.
  \end{enumerate}
\item In the unsat-core track, the \akey{check-sat} command (which is
  always issued in an unsatisfiable context) is followed by a single
  \akey{get-unsat-core} command.
\item The script may optionally contain an \akey{exit} command as its
  last command.  In the application track, this command must not be
  omitted.
\item No other commands besides the ones just mentioned may be used.
\end{enumerate}
%
The SMT-LIB format specification is available from the ``Standard''
section of the SMT-LIB website~\cite{SMT-LIB}.  Solvers will be given
formulas only from the divisions into which they have been entered.

@@

\subsection{Unsat-core track}
\label{sec:exec:unsat-core}

The unsat-core track will evaluate the capability of solvers to
generate unsatisfiable cores (for problems that are known to be
unsatisfiable).  Solvers will be measured by the smallness of the
unsatisfiable core they return.

The SMT-LIB language accommodates this functionality by providing two
features: the ability to name top-level (asserted) formulas, and the
ability to request an unsatisfiable core after a \akey{check-sat}
command returns \texttt{unsat}.  The unsatisfiable core that is
returned must consist of a list of names of formulas, in the format
prescribed by the SMT-LIB standard.

The result of a solver is considered erroneous if the response to the
\akey{check-sat} command is \texttt{sat}, or if the returned
unsatisfiable core is not well-formed (e.g., contains names of
formulas that have not been asserted before), or if the returned
unsatisfiable core is not, in fact, unsatisfiable.  The organizers
intend to check unsatisfiability by running an SMT solver on a
benchmark that asserts only those formulas in the reported core.  The
details of the checking procedure, which will be the same for all
solvers and benchmarks in the unsat-core track, are still to be
determined.

Solvers must respond to each command in the benchmark script with the
answers defined in the SMT-LIB format specification.  In particular,
solvers that respond \texttt{unknown} to the \akey{check-sat} command
must respond with an error to the following \akey{get-unsat-core}
command.

@@

\header{Benchmark sources.}
%
Benchmarks for each division will be drawn from the SMT-LIB benchmark
library.  The main track will use a subset of all
\emph{non-incremental} benchmarks; the application track will use a
subset of all \emph{incremental} benchmarks.  The unsat-core track
will use unsatisfiable main track benchmarks, modified to use named
top-level assertions.

@@

\header{Main track.} More specifically, for the main track, we have
%
\begin{itemize}
\item $e=0$, $n=0$ if the solver aborts without a response, or the
  result of the \akey{check-sat} command is \texttt{unknown},
\item $e=0$, $n=1$ if the result of the \akey{check-sat} command is
  correct,
\item $e=1$, $n=0$ if the result of the \akey{check-sat} command is
  incorrect.
\end{itemize}
%
Note that a (correct or incorrect) response is taken into
consideration even when the solver process terminates abnormally, or
does not terminate within the time limit.  Solvers should take care
not to accidentally produce output that contains \texttt{sat} or
\texttt{unsat}.

@@

\header{Unsat-core track.}
%
For the unsat-core track, we instead have $0 \leq n \leq A$, where~$A$
is the number of named top-level assertions in the benchmark, and
\begin{itemize}
\item $e=0$, $n=0$ if the solver aborts without a response, or the
  result of the \akey{check-sat} command is \texttt{unknown},
\item $e=1$, $n=0$ if the result is erroneous according to
  Section~\ref{sec:exec:unsat-core},
\item otherwise, $e=0$, and $n$ is the reduction in the number of
  formulas, i.e., $n = A$ minus the number of formula names in the
  reported unsatisfiable core.
\end{itemize}

@@

For a given competition benchmark~$b$, let~$F_b \geq 1$ be the total
number of benchmarks in $b$'s benchmark family that were used in the
competition track to which the division belongs.  We define the weight
for benchmark~$b$ as $\alpha_b = (1 + \log_e F_b) / F_b$.\footnote{See
  Section~\ref{sec:competition-wide} for a motivating discussion of
  log scaling.}  We define the \emph{normalized weight} for
benchmark~$b$ as $\alpha'_b = \alpha_b / (\sum_{b'} \alpha_{b'})$,
where the sum is over all benchmarks in the division.

For main track and unsat-core track divisions, we will separately
compute the weighted sum of all raw scores
(Section~\ref{sec:benchmark-scoring}) $$\sum_b \alpha'_b \cdot \langle
e_b, n_b, w_b, c_b\rangle$$ where the sum is over all benchmarks in
the division to assess parallel performance, and the weighted sum of
all sequential scores (Section~\ref{sec:sequential}) to assess
sequential performance.  For application track divisions, division
scores will be based on raw scores only.\footnote{Since application
  track benchmarks may be partially solved, defining a useful
  sequential score for the application track would require information
  not provided by the raw score, e.g., detailed timing information for
  each result.}
